---
title: "Named Entities"
---

Named entities are the people, places, organizations, dates, and other proper references that structure the memoir.

For this project, entity extraction is intentionally treated as a **candidate-and-review workflow**. OCR is noisy, family names are domain-specific, and the source text may be lower-case in places, so a generic named-entity model would miss important names and hallucinate others.

## How to read this page

- **Entity registry**: one row per candidate entity across the memoir.
- **Entity mentions**: chapter-level mentions of each entity.
- **Dates**: four-digit year candidates. These need review because some years are chapter-period dates while others are historical background or later reflection.
- **Review status**: starts as `candidate`. Later we will promote trusted entities to `accepted` and remove false positives.

```{r}
#| label: entities-load
#| echo: false
#| message: false
#| warning: false

library(readr)
library(dplyr)
library(DT)

if (!file.exists("data/public/entity_registry.csv") || !file.exists("data/public/entity_mentions.csv")) {
  source("R/09_extract_named_entities.R")
}

registry <- readr::read_csv("data/public/entity_registry.csv", show_col_types = FALSE)
mentions <- readr::read_csv("data/public/entity_mentions.csv", show_col_types = FALSE)
summary <- readr::read_csv("data/public/entity_summary.csv", show_col_types = FALSE)
```

## Entity counts by type

```{r}
#| echo: false
DT::datatable(summary, rownames = FALSE, options = list(pageLength = 10))
```

## Top entities

```{r}
#| echo: false
registry |>
  arrange(desc(total_count)) |>
  select(entity_type, canonical_entity, total_count, chapters, extraction_methods, review_status) |>
  head(100) |>
  DT::datatable(rownames = FALSE, options = list(pageLength = 25, scrollX = TRUE))
```

## Chapter-level mentions

```{r}
#| echo: false
mentions |>
  select(chapter, chapter_title, entity_type, canonical_entity, count, extraction_method, review_status) |>
  arrange(chapter, entity_type, desc(count), canonical_entity) |>
  DT::datatable(rownames = FALSE, options = list(pageLength = 25, scrollX = TRUE))
```

## What comes next

The next improvement is to manually edit `data/entity_seeds.csv` and add names, places, institutions, and events that the automatic pass missed. After that, we can promote reviewed rows into a durable entity registry for the graph database and chatbot retrieval layer.
