---
title: "Uncle Virgil Memoir Analysis"
subtitle: "A Quarto rebuild of the biography text-analysis and chatbot project"
---

```{r}
#| label: load-index-data
#| echo: false
library(readr)
library(dplyr)
metadata <- read_csv("data/chapter_metadata.csv", show_col_types = FALSE)
stats <- read_csv("data/processed/chapter_stats.csv", show_col_types = FALSE)
```

This site analyzes the Uncle Virgil memoir as a digital humanities corpus. The first rebuild focuses on four things:

1. preserving the cleaned OCR corpus,
2. structuring the memoir by chapter,
3. rebuilding text analysis in reproducible R scripts,
4. adding search and a future-ready chatbot interface.

## Current corpus status

```{r}
#| echo: false
stats %>%
  filter(chapter != 12) %>%
  summarise(
    chapters = n(),
    total_words = sum(word_count, na.rm = TRUE),
    total_characters = sum(char_count, na.rm = TRUE)
  )
```

## Chapter word counts

```{r}
#| echo: false
library(ggplot2)
stats %>%
  filter(chapter != 12) %>%
  ggplot(aes(x = reorder(chapter_title, chapter), y = word_count)) +
  geom_col() +
  coord_flip() +
  labs(x = NULL, y = "Words", title = "Chapter length by word count") +
  theme_minimal()
```

## Rebuild roadmap

The Shiny app has been replaced conceptually with a static Quarto front end. The model-driven bot should sit behind a backend API rather than inside the static site.

```text
Quarto static site
  -> local search index
  -> optional AWS API Gateway endpoint
  -> Lambda retrieval function
  -> future model layer
```
