---
title: "Sentiment Analysis"
page-layout: full
toc: true
---

This page shows two complementary sentiment views:

1. **Sentiment by chapter** — a chapter-level summary of positive and negative language.
2. **Sentiment within each chapter** — passage-level sentiment so you can see where tone rises and falls inside a chapter.

The current method uses the `tidytext` Bing sentiment lexicon. This is a transparent dictionary-based method, not a model-generated interpretation. Treat it as a reading aid, not a final literary judgment.

```{r}
#| label: sentiment-load
#| message: false
#| warning: false
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(stringr)

chapter_sentiment <- read_csv("data/public/sentiment_by_chapter.csv", show_col_types = FALSE)
passage_sentiment <- read_csv("data/public/sentiment_within_chapter.csv", show_col_types = FALSE)
passage_summary <- read_csv("data/public/sentiment_passage_summary.csv", show_col_types = FALSE)
```

## Sentiment by chapter

This view compares chapters after adjusting net sentiment by chapter length. Positive values indicate more positive sentiment words per 1,000 words; negative values indicate more negative sentiment words per 1,000 words.

```{r}
#| label: sentiment-by-chapter-plot
#| fig-width: 12
#| fig-height: 7
chapter_sentiment |>
  mutate(chapter_label = factor(chapter_label, levels = rev(chapter_label))) |>
  ggplot(aes(x = chapter_label, y = net_per_1000_words)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  labs(
    title = "Net sentiment by chapter",
    subtitle = "Positive minus negative sentiment words, normalized per 1,000 words",
    x = NULL,
    y = "Net sentiment per 1,000 words"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
#| label: positive-negative-by-chapter
#| fig-width: 12
#| fig-height: 7
chapter_sentiment |>
  select(chapter, chapter_label, positive_count, negative_count) |>
  pivot_longer(c(positive_count, negative_count), names_to = "sentiment", values_to = "count") |>
  mutate(
    sentiment = recode(sentiment, positive_count = "Positive", negative_count = "Negative"),
    chapter_label = factor(chapter_label, levels = rev(unique(chapter_sentiment$chapter_label)))
  ) |>
  ggplot(aes(x = chapter_label, y = count, fill = sentiment)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Positive and negative sentiment word counts by chapter",
    subtitle = "Raw counts are useful, but longer chapters naturally have more sentiment words",
    x = NULL,
    y = "Sentiment word count",
    fill = NULL
  ) +
  theme_minimal(base_size = 13)
```

```{r}
#| label: sentiment-by-chapter-table
chapter_sentiment |>
  transmute(
    Chapter = chapter,
    Title = chapter_title,
    `Word count` = word_count,
    Positive = positive_count,
    Negative = negative_count,
    `Net sentiment` = net_sentiment,
    `Net per 1,000 words` = net_per_1000_words,
    `Sentiment words` = sentiment_word_count
  ) |>
  knitr::kable()
```

## Sentiment within each chapter

This view uses the passage index. Each point is one passage. This is closer to what the original app was trying to show: not just whether a whole chapter trends positive or negative, but where the tonal shifts occur inside the chapter.

```{r}
#| label: within-chapter-sentiment-plot
#| fig-width: 14
#| fig-height: 14
passage_sentiment |>
  ggplot(aes(x = passage_number, y = net_per_1000_words, group = chapter)) +
  geom_hline(yintercept = 0, linewidth = 0.25) +
  geom_line(linewidth = 0.35) +
  geom_point(size = 1.4) +
  facet_wrap(~ chapter_label, scales = "free_x") +
  labs(
    title = "Sentiment shifts within each chapter",
    subtitle = "Each point is a passage from the passage index",
    x = "Passage number",
    y = "Net sentiment per 1,000 words"
  ) +
  theme_minimal(base_size = 12)
```

## Strongest passage-level shifts

These are useful review targets. They show where the lexicon method finds the strongest positive and negative passages. Read these against the original text before making interpretive claims.

```{r}
#| label: strongest-passages-table
strongest_positive <- passage_sentiment |>
  filter(sentiment_word_count >= 3) |>
  arrange(desc(net_per_1000_words)) |>
  slice_head(n = 10) |>
  mutate(Direction = "Most positive")

strongest_negative <- passage_sentiment |>
  filter(sentiment_word_count >= 3) |>
  arrange(net_per_1000_words) |>
  slice_head(n = 10) |>
  mutate(Direction = "Most negative")

bind_rows(strongest_positive, strongest_negative) |>
  transmute(
    Direction,
    Citation = citation_label,
    `Net per 1,000 words` = net_per_1000_words,
    Positive = positive_count,
    Negative = negative_count,
    Years = year_candidates,
    Entities = str_trunc(entities_mentioned, 100),
    Preview = text_preview
  ) |>
  knitr::kable()
```

## Passage sentiment table

```{r}
#| label: passage-sentiment-table
passage_sentiment |>
  transmute(
    Citation = citation_label,
    `Word count` = word_count,
    Positive = positive_count,
    Negative = negative_count,
    `Net sentiment` = net_sentiment,
    `Net per 1,000 words` = net_per_1000_words,
    Direction = sentiment_direction,
    Years = year_candidates,
    Entities = str_trunc(entities_mentioned, 120)
  ) |>
  head(60) |>
  knitr::kable()
```
