---
title: "Timeline"
---

The timeline below uses **curated chapter date ranges** from `data/chapter_metadata.csv`, not raw year extraction alone.

That distinction matters. The OCR/RDS rebuild now recovers years, but a memoir chapter can mention years outside the life period being narrated. For example, a childhood chapter may mention a later anniversary, a publication date, or a retrospective comment. Those years are useful evidence, but they should not automatically define the chapter's time range.

```{r}
#| label: timeline-load
#| echo: false
#| message: false
#| warning: false

library(readr)
library(dplyr)
library(ggplot2)
library(knitr)

if (!file.exists("data/public/chapter_timeline.csv")) {
  source("R/08_build_timeline_table.R")
}

timeline <- readr::read_csv("data/public/chapter_timeline.csv", show_col_types = FALSE) |>
  filter(chapter != 12)
```

## Chapter date ranges

```{r}
#| echo: false

ggplot(timeline, aes(y = reorder(timeline_label, desc(chapter)))) +
  geom_segment(aes(x = approx_start_year, xend = approx_end_year, yend = reorder(timeline_label, desc(chapter))), linewidth = 4) +
  geom_point(aes(x = approx_start_year), size = 2) +
  geom_point(aes(x = approx_end_year), size = 2) +
  labs(
    x = "Approximate year range",
    y = NULL,
    title = "Curated memoir timeline by chapter",
    subtitle = "Ranges are approximate and should be reviewed against the scans."
  ) +
  theme_minimal()
```

## How to read this

- **Start/end years** are human-curated approximations based on the chapter's main narrative period.
- **Year candidates** are machine-detected years found inside the OCR text.
- A candidate year outside the curated range is not automatically wrong; it may be historical background, a retrospective reference, or OCR noise.
- Chapter 12 is excluded from the chart because it is the synthetic full-text aggregate.

```{r}
#| echo: false

timeline |>
  select(
    chapter,
    chapter_title,
    approx_start_year,
    approx_end_year,
    life_stage,
    primary_locations,
    word_count,
    year_count,
    year_candidates
  ) |>
  knitr::kable()
```
